### 2021-10-08 plos_latex_template.tex
ast: financial rule + general reading, all WIP.

The authors of this article participated in the organization team of the annual user conference of the R Project for Statistical Computing, held in July 2021. useR! conferences are non-profit events, organized by volunteers from the R community and arranged by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we share a few lessons learned in the process. We streamline our most important learnings in 10 simple rules to host an inclusive conference. These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.

@@

Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields.
Conversely, efforts to address systematic discrimination in conferences and make them inclusive spaces can have a meaningful impact in our communities and the lives and careers of their members.

@@

The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, that coordinates with the local organizers of conferences, or even the local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, truly inclusive and accessible experience for everyone. 
Here, we share the lessons learned within the past year of organizing this global useR!2021, summarized as 10 simple rules towards an inclusive conference.
% Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing and the whole structure

@@

Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. 
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems--including conferences--are consciously and unconsciously built.  
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have excluded systematically whole groups of people from academia and from scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

To be proactive allies in our conferences and communities, we should start by examining our own privileges (unearned advantages given by society to some people but not all).
Privilege is largely invisible to those who have it. And we will not be able to help end discrimination and oppression if we do not recognize our privileges first and the injustices that come with them.
Once we understand our own privileges, we can begin addressing them on an individual, collective, and institutional basis.
Recognizing our privileges--particularly in our field and in our scientific or professional community--will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include. 
Depending on the field, region or community, it could be Black, LGBTQIA+, Muslim, or others. 
Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.

@@

Paying attention to some dimensions of diversity while neglecting others may have unintended, even harmful consequences. Lack of representation, unwelcoming--or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members. A negative conference experience can divert career paths, affect lives, and exclude people from some fields \cite{biggsAcademicConferenceChilly2018}. 
%Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.
% Rocío: Andrea, could the newbies part go in the networking (box)? I don't see how to make it fit here. 
%i took the networking part to expand communication att.ast we can switch back afterwards

@@

The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct (CoC). We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct team work.

%Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the outreach and communication strategy see Rule \ref{rule_communication}. However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these broader goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
% Rocío: This could be moved to communication (Batool, Andrea)

@@

A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team. Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity. Particular attention should be paid to the usually marginalized groups (see \textbf{Rule \ref{rule_diversity}}). To ensure a deeper understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. If you already have an organizing team, check for gaps in its composition. 
%Rocío: I changed "biases" for "gaps". I'm sure other statisticians would've found the phrasing weird. To really understand what is meant by biases, the reader would have to go to rule unbias, and rules should be self-contained.
% Rocío: Idea of this paragraph: build a diverse team

Gathering a diverse team will only work if there is real inclusion. People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity. This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups. A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear. 
%liz, please check if this paragraph connects well with the whole rule. i am using nothing about us without us but also talking about all marginalized groups, in the context of building a team. att: andrea
% Liz: Yes!

@@

%This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees. This gave a sense of variety and an appeal of “at least one networking session for everyone”.  (Marcela)
%ast 17sep: i think we can talk here about planning neworking here, but also: the whole paragraph applies to team dynamics in general. Sara: maybe you can use this as we spoke relating team work here. And I think the working together part is not here but the following paragraph 
%sara: The last sentence in the paragraph sounds odd to me, but maybe I'm missing something. Andrea, I added a few sentences to the following paragraph regarding work dynamics
% Rocío: the rule is getting too long and, since this is focused on networking, I'd prefer it to go to a networking section.

% Liz: I know this section is getting long, and maybe this is too blunt, but: 
% People who have experience with exclusion have the social and technical 
% expertise about removing barriers to inclusion.Actively ask where the barriers are and how to mitigate them.

The following advice should apply to any kind of team, but are specially relevant when working with people from marginalized groups. First, rethink power relationships inside the group and share power. Be committed to examine power relations inside your organizing team and challenge it during all the process. % Rocío: Could we explain power relationship / share power ideas in more simple terms?
Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer leading positions, and let people from privileged groups step down. 
Build an environment in which every person can express their position and give prevalence to the positions from systematically excluded groups.
% Liz: Use of "voice" in this way makes deaf people feel excluded. Maybe it could be something like "Build an environment in which every person has input..." Or anything as long as it doesn't suggest you have to have a voice to participate.
%i rephrased slightly, avoiding "support" it's not a question of privileged people giving support but of positions being given importance
Offer support and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See \textbf{Rule \ref{rule_language}}), or others. You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership. Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}). Make work visible, discuss collectively what should be done (and why), divide and account work. Being transparent and sharing the information is key to make work visible and collective. %sara: this relates to make labor visible from DF, maybe we can cite them here too. %Rocío: sure.
% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.

@@

When choosing or inviting people as keynote speakers, program committee, session chairs and other spotlight roles, it is likely that there will not be much diversity in the first set of names. Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. Rather than deter us, this implicit and systemic bias should encourage us to look further to find great people that are not routinely in the spotlight. 
Ensuring diversity in each of these roles needs to be a deliberate process. We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.

Make sure that every selection committee--the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees--are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has implicit biases, to recognize them, and try to counteract them. 
An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles. The regional and local communities in your field are also good sources to tap into. 
For example, for useR!2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
When organizing an international academic conference, you may reach local associations, student organizations, or other Early Career Researchers groups or communities of practice: ask for their barriers for participation, and explicitly invite them to contribute to the conference. % Rocío: ask about their requirements and barriers sounds a bit confusing.
% Rocío: Not everyone knows what a community of practice is. Could we add a definition or use other words? Definition: "Communities of practice are groups of people who share a concern, a set of problems, or a passion about a topic, and who deepen their knowledge and expertise in this area by interacting on an ongoing basis". (Wenger, E., McDermott, R. A., & Snyder, W. (2002). Cultivating Communities of Practice: A Guide to Managing Knowledge. Harvard Business Press.)
Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged. Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers and people from minoritized groups, fostering collaboration and building their skills.
% Liz: maybe consider a broader group than "early career researchers." One problem is that marginalized people have many fewer opportunities, and as they age that attenuation of opportunities compounds. Someone can be in the margins even years after finishing grad school or being a postdoc (not to mention that people without degrees could have contributions.

Furthermore, consider bringing to the spotlight the people that have contributed to your field in more collaborative ways \cite{cheng2020x+}. You can, for instance, reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), to acknowledge being mindful of inclusiveness. Community building, for instance, is challenging and usually unrewarded when compared to publications or software development \cite{acionWhyChooseCommunity2020}; give the people who deliberately took the time to work on their communities the recognition they deserve. Defy the stereotypical criteria for success by acknowledging these community practices. 
Finally, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to Diversity, Inclusion, and Accessibility. Recognize their areas of expertise and respect their will regarding participation (or not) in community-building events.
% Rocío: Since the system usually rewards the most competitive people, appealing cvs and so on, I think it's worth it to explain that rewarding collaborative work is also part of unbiasing; not only because it's mostly done by women, people of color, etc., but because this kind of amazing work is usually neglected in our conferences. It may not be obvious to everyone that the content of this paragraph is not only "bring these people to the spotlight", but that it also means unbiasing the conference. I don't think I'll have time to add lines about this in the beginning of the paragraph so if anybody could do that, it would be great.

% Liz: I might be able to, but I'm not sure where you want the additional text; where the beginning of the paragraph is because of how this is formatted. I couldn't agree more that being truly inclusive means not restricting pepole's role. Every subteam should be diverse, in addition to the whole organizing team. It's not just about representation but needs to encompass integration across the whole event.

@@

\section{Allocate financial resources to support the presence of diverse people in every space of the conference
}
\label{rule_financial}

\subsection*{Inclusion comes with a cost}

Many inclusive practices come with a cost and it is important to add them to the budget when planning the conference. 
Good quality and professional captioning, interpreters or translators, accessible and screen-reader friendly registration tools--and chat platform if the conference is online--, paying your organizing team (see \textbf{Rule \ref{rule_organizing_team}}), and code of conduct training, are of the some expenses that you may have to include in your budget. 
Budgets are limited, so it is important to estimate costs in advance and define your priorities.

\subsection*{Set fair registration rates and provide financial support when needed}

Conferences, even virtual ones, should have registration fees for two primary reasons. The first one is that preparing the conference costs money and this is likely to translate to the participation fee. The second reason has to do with psychology: people value more the things they pay for, and there is a lower attendance rate for free events than for events with registration costs \cite{eventbrite_ultimate_2017}.

On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates. Usually, there is a higher fee for people from the industry than for academia. A lower fee for non-profit organizations, government employees, or freelancers should also be considered. It is important to include discounts for students to encourage their participation \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}. Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, they should be allowed to locate themselves in a category they consider affordable (i.e. student, researcher or `pay what you can'). To fix the registration costs, you can take into account the cost of living in each country using conversion factors (e.g. from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}). Consider that, while it is traditional for employers to provide--at least some--conference support in the Global North, this is not the case everywhere.

Offering scholarships or grants to attend the conference are a common way to boost participation from diverse groups, and some conferences ask for cover letters or applications  but they may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them. 
Also, the process for applicants should be simplified. Applying for loans, grants, and scholarships may be an emotionally demanding task. Do not complicate the attendees' lives by asking for long essays to convince a committee that they deserve your support. For this stage of the process, a certain degree of trust goes a long way-- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility. 
Only if the applicants for fee waivers are numerous and your budget cannot allow for that many fee waivers, you could ask for a short application. You could try to reduce this risk by finding enough sponsoring money so that you do not depend on registration fees to pay for your conference.

If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees and visa financial and administrative support (if in-person), or internet connection services (if virtual). If the conference is virtual, you may offer technical support so that attendees can record their videos, include captions, or design accessible presentations.

% If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or terms that single out the recipients (e.g. `grant recipient' or `diversity scholar', respectively). Do not call it a scholarship or a grant. Ultimately, it is a fee waiver or a discount for the people who need additional financial support. 
% Rocío: I removed the part of naming the scholarship/grant/waiver, because if it goes somewhere it should be communication. Andrea, Batool, please see if it fits there.

>>>>

The authors of this article participated in the organization team of the annual user conference of the R Project for Statistical Computing, held in July 2021. useR! conferences are non-profit events organized by volunteers from the R community and arranged by the R Foundation. The conference attracts a broad range of participants from academia, industry, government, and the non-profit sector. For 2021, we aimed to build a high-quality virtual and explicitly global conference in a kind, inclusive, accessible, and welcoming environment for everyone. 
In this article, we share a few lessons learned in the process. We streamline our most important learnings in 10 simple rules to host an inclusive conference. These rules apply equally to academic, industry, or mixed conferences; the rules are inspired by a global experience but also apply at the regional or local level.

@@

Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
However, they are likely to reproduce the systematic discrimination occurring in other spaces in our fields and become [bad experiences]. 
Efforts to address systematic discrimination in conferences and make them inclusive spaces can have a meaningful impact on our communities and the lives and careers of their members.

@@

The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, that coordinates with the local organizers of conferences, or even the local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, truly inclusive, and accessible experience for everyone. 
Here, we share the lessons learned within the past year of organizing this global useR! 2021, summarized as ten simple rules towards an inclusive conference.
We have organized these rules in three groups, because they refer to the foundations (Rule 1 and 2), the actions regarding people (Rules 3 to 5), the actions regarding conference design (Rules 6 to 9), and a general synthesis %ast this is an attempt, but i am not a fan of rule 10 alone at the end. actually the whole classification when written seems off to me (right now, this might change)
% Rocío: Could add a line introducing Fig. in https://docs.google.com/drawings/d/1iS1pLc9OldLMe_jhNT6OGIbT9s5KvlHpUoDd19vHkCY/edit?usp=sharing and the whole structure [x]

@@

Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. 
Human diversity should be celebrated and respected in every way. Nonetheless, we live in a world with implicit hierarchies along these axes. Some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems--including conferences--are consciously and unconsciously built.  
While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

To be proactive allies in our conferences and communities, we should start by examining our own privileges (unearned advantages given by society to some people but not all).
Privilege is largely invisible to those who have it. And we will not be able to help end discrimination and oppression if we do not recognize our privileges first and the injustices that come with them.
Once we understand our own privileges, we can begin addressing them on an individual, collective, and institutional basis.
Recognizing our privileges--particularly in our field and in our scientific or professional community--will help identify which subgroups have been the most excluded or discriminated against. 
These are the groups we need to make more effort to include. 
Depending on the field, region, or community, it could be Black People, LGBTQIA+, Muslim, or others. 
%ast: ö check phrasing, add "people"
Intersectionality, or the complex, cumulative way in which the effects of multiple forms of discrimination combine or intersect, should be taken into account.

@@

Paying attention to some dimensions of diversity while neglecting others may have unintended, even harmful consequences. Lack of representation, unwelcoming--or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members. A negative conference experience can divert career paths, affect lives, and exclude people from some fields \cite{biggsAcademicConferenceChilly2018}. 
%ast maybe this last sentence could go in the introduction because it's general i marked [bad experience] in the intro

@@

The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct (CoC). We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for the Code of Conduct teamwork.

@@

A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team. Build a team with people from different regions, genders, races, socioeconomic statuses, and other aspects of diversity. 
Go beyond balancing all genders in this effort, and pay attention to other marginalized groups (see \textbf{Rule \ref{rule_diversity}}). To ensure a deep understanding and smooth communication with different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. If you already have an organizing team, check for gaps in its composition.

%ast: I added race -even if race is a construct- because focus is given to gender and the spaces tend to be predominantly white and US/Europe-centric. In other parts of the text I was using racialized but here we can keep it simple, i think. another thing: i avoided "gender balancing" because idk if that comes across as too binary, even if what is done is adding cis women to the team

% Rocío: Idea of this paragraph: build a diverse team

Gathering a diverse team will only work if there is real inclusion. People with disabilities often say: `Nothing about us without us'; the same holds for other dimensions of diversity. This means that the actual life experiences, expertise, and insights from people in marginalized groups cannot be replaced by good intentions from people outside these groups [cite design justice]. A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear.

@@

%This is specially important for networking spaces, where receiving ideas, direct feedback, and spaces to co-lead from a wide variety of community members can be the key to organize meaningful sessions for all the conferences attendees. This gave a sense of variety and an appeal of "at least one networking session for everyone".  (Marcela)
%ast 17sep: i think we can talk here about planning neworking here, but also: the whole paragraph applies to team dynamics in general. Sara: maybe you can use this as we spoke relating team work here. And I think the working together part is not here but the following paragraph 
%sara: The last sentence in the paragraph sounds odd to me, but maybe I'm missing something. Andrea, I added a few sentences to the following paragraph regarding work dynamics
% Rocío: the rule is getting too long and, since this is focused on networking, I'd prefer it to go to a networking section.

% Liz: I know this section is getting long, and maybe this is too blunt, but: 
% People who have experience with exclusion have the social and technical 
% expertise about removing barriers to inclusion. Actively ask where the barriers are and how to mitigate them.

The following advice should apply to any kind of team, but is specially relevant when working with people from marginalized groups. First, rethink power relationships inside the group and share power. Be committed to examine power relations inside your organizing team and challenge it during all the process. % Rocío: Could we explain power relationship / share power ideas in more simple terms?
Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer leading positions, and let people from privileged groups step down. 
Build an environment in which every person can express their position and give prevalence to the positions from systematically excluded groups.
% Liz: Use of "voice" in this way makes deaf people feel excluded. Maybe it could be something like "Build an environment in which every person has input..." Or anything as long as it doesn't suggest you have to have a voice to participate.
%ast: i rephrased slightly, avoiding "support" it's not a question of privileged people giving support but of positions being given importance
Offer support and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See \textbf{Rule \ref{rule_language}}), or others. You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership. Splitting the workload and responsibilities should not be done by putting care-taking labors--community building, meeting organization and note-taking, conversations with potential partners--on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see \textbf{Rule \ref{rule_unbias}}). Make work visible, discuss collectively what should be done (and why), divide and account work. Being transparent and sharing the information is key to make work visible and collective. %sara: this relates to make labor visible from DF, maybe we can cite them here too. %Rocío: sure.
% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work.

@@

%idea1: our lists are biased

When choosing or inviting people as keynote speakers, program committee, session chairs, and other spotlight roles, it is likely that there will not be much diversity in the first set of names. Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. Rather than deter us, this implicit and systemic bias should encourage us to look further to find great people that are not routinely in the spotlight. 
Ensuring diversity in each of these roles needs to be a deliberate process. We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.

Make sure that every selection committee--the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees--are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has implicit biases, to recognize them, and try to counteract them. 
An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles. The regional and local communities in your field are also good sources to tap into. 
For example, for useR!2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
When organizing an international academic conference, you may reach local associations, student organizations, or other Early Career Researchers groups or communities of practice: ask for their barriers for participation, and explicitly invite them to contribute to the conference. 
% Rocío: Not everyone knows what a community of practice is. Could we add a definition or use other words? Definition: "Communities of practice are groups of people who share a concern, a set of problems, or a passion about a topic, and who deepen their knowledge and expertise in this area by interacting on an ongoing basis". (Wenger, E., McDermott, R. A., & Snyder, W. (2002). Cultivating Communities of Practice: A Guide to Managing Knowledge. Harvard Business Press.)
Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged. Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers and people from minoritized groups, fostering collaboration and building their skills.
% Liz: maybe consider a broader group than "early career researchers." One problem is that marginalized people have many fewer opportunities, and as they age that attenuation of opportunities compounds. Someone can be in the margins even years after finishing grad school or being a postdoc (not to mention that people without degrees could have contributions.

Furthermore, consider bringing to the spotlight the people that have contributed to your field in more collaborative ways \cite{cheng2020x+}. You can, for instance, reframe the awards ceremony to acknowledge those who prepared accessible slides and presentations (see Rule \ref{rule_accessibility}), for being mindful of inclusiveness. Community building, for instance, is challenging and usually unrewarded when compared to publications or software development \cite{acionWhyChooseCommunity2020}; give the people who deliberately took the time to work on their communities the recognition they deserve. Defy the stereotypical criteria for success by acknowledging these community practices. 
Finally, do not restrict people from marginalized groups or community-builders to talk or work only in issues related to Diversity, Inclusion, and Accessibility. Recognize their areas of expertise and respect their will regarding participation (or not) in community-building events.
% Rocío: Since the system usually rewards the most competitive people, appealing cvs and so on, I think it's worth it to explain that rewarding collaborative work is also part of unbiasing; not only because it's mostly done by women, people of color, etc., but because this kind of amazing work is usually neglected in our conferences. It may not be obvious to everyone that the content of this paragraph is not only "bring these people to the spotlight", but that it also means unbiasing the conference. I don't think I'll have time to add lines about this in the beginning of the paragraph so if anybody could do that, it would be great.

% Liz: I might be able to, but I'm not sure where you want the additional text; where the beginning of the paragraph is because of how this is formatted. I couldn't agree more that being truly inclusive means not restricting people's role. Every subteam should be diverse, in addition to the whole organizing team. It's not just about representation but needs to encompass integration across the whole event.

% ast: i agree a lot with Rocío here, the change in what is given importance is a key part of the unbiasing. i had the impression that the part of not restricting people's role goes above, in team rule.

@@

\section{Allocate financial resources to support your conference goals}
\label{rule_financial}

%Inclusion comes with a cost

Conference budgets are limited and rely mostly on sponsorship and attendance fees. Some expenses are more or less fixed but allocation of resources has to be intentional to support the goals of the conference. It is important to estimate the costs for these inclusive practices and define your priorities in advance. This includes paying at least part of your organizing team (see \textbf{Rule \ref{rule_organizing_team}}), training (e.g., code of conduct, active bystandership training, \textbf{Rule \ref{rule_inclusion}}), and accessibility practices (\textbf{Rule \ref{rule_accessibility}}). It also includes additional support for attendees: child care support, transportation fees, visa-related support (if in-person), internet connection services (for the virtual component). While some of these items are being implemented recently, others can encounter a certain degree of resistance, such as paying the organizing team. Don't be afraid to innovate and resist the institutional inertia. "This has always been like that" will not take the conference towards structural change.
%should we talk about the pandemic? something like: 
With the COVID-19 pandemic and the shift to online (or hybrid) events, some expenses disappeared, leaving space for rearrangements in this sense.
Depending on registration fees to support large budget items is not at all desirable, instead sponsors can be asked to support specific items in this list.

%Set fair registration rates
Registration costs are one of the largest barriers for conference attendance, and, if we are aiming for inclusiveness and representation, the socioeconomic context of participants, their country of origin, and their career status should be taken into account when determining the registration rates. Usually, there is a higher fee for people from the industry than for academia. A lower fee for non-profit organizations, government employees, or freelancers should also be considered. It is important to include discounts for students \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012} and other early career stages, such as postdoc or trainee. The definition of such positions and the associated work conditions vary for each country. Consider that, while it is traditional for employers to provide--at least some--conference support in the Global North, this is not the case everywhere. In addition to this, low rates in strong currencies can represent significant amounts of money in other currencies. You can adjust the registration fee by the cost of living in each country using conversion factors (e.g. from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}, check https://spcanelon.github.io/useR2021-cost-conversion-tool/ for an example). In general, people should be allowed to locate themselves in a category they consider affordable, even with the possibility of a "pay what you can" approach. Resources permitting, you can aim to have a conference with no registration costs, but bear in mind that free events have a lower attendance rate than events with registration costs \cite{eventbrite_ultimate_2017}. Offering fee waivers to part of the participants is also a good option, but even then, other costs can be prohibitive.

Offering scholarships or grants to attend the conference is a common way to boost participation from different groups, and some conferences ask for cover letters or applications to them. These applications may be counterproductive, if they attract people that do not need the money but apply to enrich their CVs, and can filter out people who do not feel entitled to earn them.  
Applying for loans, grants, and scholarships may be an emotionally demanding task. Do not complicate the attendees' lives by asking for long essays to convince a committee that they deserve your support. For this stage of the process, a certain degree of trust goes a long way-- when someone says they need the support, it is best to give them the waiver rather than second guess their eligibility. If the applicants for fee waivers are numerous and your budget cannot allow for that many fee waivers, you could ask for a short application.
